---
title: "Collective risk"
author: "Ignacio Gómez, Gonzalo Pato, Gonzalo Prats"
institute: "Simulation in Prob and Stats BSc AMC at UC3M"
date: "April 2023"
output:
  pdf_document:
    toc: true
    toc_depth: 1
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.fullwidth=TRUE, fig.height=3)
library(ggplot2)
library(cowplot)
```

# Statement

The clients of an insurance company generate claims according to independent (homogeneous) Poisson processes with common rate $\lambda$.
The claims are independent and the amount of each of them follows a distribution with cdf $F(x) = 1-(100/x)^{2.5}$ if $x > 100$.
New clients enroll the company according to another (homogeneous) Poisson process with rate $\nu$ and the time they stay in the company follows an exponential distribution with rate $\mu$.
Each policyholder pays a fixed amount $a$ every year to the company.
If the policyholder stays in the company only during a fraction of the year, she will pay the corresponding fraction of $a$ which is assumed to be paid on a continuous manner.

If the initial capital is $c_0$ and the initial number of clients is $n_0$, we want to compute the probability that the capital of the company remains positive during some given time $t_l$.

# Introduction

Each policyholder has a fixed amount $a$ to pay every year, so the capital of the company will increase by $a$ times the number of clients every fraction of year.

On the other hand, they also hold the right to claim the premium on the policy, which happens according to a Poisson process with common rate $\lambda$.
The amount of is this premium is distributed according to a Pareto distribution with parameters $\alpha = 2.5$ and $\beta = 100$.

The clients of the company can leave the company at any time, and the time they stay in the company follows an exponential distribution with rate $\mu$.

Clients come according to an independent (homogeneous) Poisson processes with rate $\nu$, and they stay according to an exponentially distributed time with rate $\mu$.

So in general, the capital of the company at any time $t$ will be: 

$$C(t) = c_0 + at(n_0+N_A(t)-N_D(t)) - \sum_{j=1}^{N_C(t)} X_j$$

where $N_A(t)$ is the number of clients that arrive by time $t$, $N_D(t)$ is the number of clients that leave by time $t$, $N_C(t)$ is the number of claims that arrive by time $t$, $X_j$ is the amount of the $j$-th claim, and $n(t)$ is the number of clients at time $t$.

# The project 

Project development including the code and emphasizing the main difficulties and most important parts.

First, to simulate the amount of the claims we will define a function with the inverse of the cdf of the Pareto distribution. Mathematically, this function is:

$$F^{-1}(x) = \frac{\beta}{\sqrt[\alpha]{(1-x)}} = \frac{100}{\sqrt[2.5]{(1-x)}}$$

```{r inverseclaim, echo=TRUE, message=FALSE, warning=FALSE}
inverse.claim <- function(x) {
  return(max((100 / (1 - x)^(1 / 2.5)), 0))
}
```

The algorithm we will follow is, for each path:

1.  Initialize variables $t=0, n=n0, c=c0$.

2.  Compute the time of the first event: $dt = rexp(1, lambda*n + mu*n + nu)$

3.  Update the time: $t = t + dt$

4. While $t < tl$:
    
    4.1. Update capital with the proportional part of the payments: $c = c + n*a*dt$

    4.2. Decide the type of event and update in each case:
    
    <ul>
        <li>&nbsp;&nbsp;&nbsp;&nbsp;4.2.1. If $u < \lambda*n/(\lambda*n + \mu*n + \nu)$, claim event so update capital $= c - X_j$</li>

        <li>&nbsp;&nbsp;&nbsp;&nbsp;4.2.2. If $u < (\lambda*n + \mu*n)/(\lambda*n + \mu*n + \nu)$, departure event so $n=n-1$</li>

        <li>&nbsp;&nbsp;&nbsp;&nbsp;4.2.3. Else, enrollment event, so $n=n+1$</li>
    </ul>
        
    4.3.  Compute the time of the next event: $dt = rexp(1, lambda*n + mu*n + nu)$

    4.4.  Update the time: $t = t + dt$

Thus for one simulation we have:

```{r simfunction}
simulate_one_insurance <- function(c0, n0, a, tl, lambda, mu, nu) {
  # Initialize variables
  t <- 0
  n <- n0
  c <- c0
  # Update capital with annuity payments
  dt <- rexp(1, lambda * n + mu * n + nu)
  t <- t + dt
  while (t < tl) {
    # Update the capital with the proportional part of the payments
    c <- c + n * a * dt
    # Decide the type of event
    u <- runif(1)
    if (u < lambda * n / (lambda * n + mu * n + nu)) {
      # Claim event
      claim_amount <- inverse.claim(runif(1))
      c <- c - claim_amount
    } else if (u < (lambda * n + mu * n) / (lambda * n + mu * n + nu)) {
      # Departure event
      n <- max(n - 1, 0)
    } else {
      # Enrollment event
      n <- n + 1
    }
    # Stop if capital becomes negative
    if (c < 0) break
    # Time of next event
    dt <- rexp(1, lambda * n + mu * n + nu)
    t <- t + dt
  }
  return(c)
}
```

And for multiple: 
```{r siminsurance}
simulate_insurance <- function(c0, n0, a, tl, lambda, mu, nu, MC) {
  final_capitals <- vector("numeric", MC)
  # Simulate MC paths
  final_capitals <- replicate(MC, simulate_one_insurance(c0, n0, a, tl, lambda, mu, nu))
  # Compute and return outputs
  fraction <- sum(final_capitals > 0) / MC
  pos_fc <- final_capitals[final_capitals > 0]
  mean_final_capitals <- mean(pos_fc)
  sd_final_capitals <- sd(pos_fc)
  p1 <- ggplot() + 
        geom_line(mapping= aes(x = 1:length(pos_fc), 
                               y = cumsum(pos_fc)/(1:length(pos_fc)))) +
        labs(x = "Number of simulations", y = "Mean of final capitals",
               title = "Convergence of the mean of the final capitals")
  plot(p1)
  p2 <- ggplot() + 
        geom_histogram(mapping= aes(x = pos_fc, y=after_stat(density)),
                       binwidth=5000, color = "black", fill = "white") + 
        stat_function(fun=dnorm, args=list(mean=mean_final_capitals, sd = sd_final_capitals), 
                      color = "red") +
        labs(x = "Final capitals", y = "Density", 
               title = "Density of the final capitals")
  plot(p2)
  return(list(fraction = fraction, mean_final_capital = mean_final_capitals,
              sd_final_capital = sd_final_capitals))
}
```

# Results 

```{r funsim}
set.seed(42)
res <- simulate_insurance(c0 = 1000, n0 = 100, a = 100, tl = 100, 
                          lambda = 0.1, mu = 0.1, nu = 0.1, MC = 10000)
res
```

Numerical results, quality assessment of the approximations, and time efficiency of the algorithms.

# Conclusions

About the results, how the difficulties were solved, and possible alternative approaches. Keep the focus, the conclusions must be as brief as possible.

# References

Including textbooks, webpages, and class notes.
