---
title: "Unit3-DiscreteEvents"
author: "Gonzalo Pato Montemayor"
date: "27-02-2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(dplyr)
library(randtests)
library(EnvStats)
library(ggplot2)
set.seed(42)
```

## Exercise 1

(a) The distribution of the total number of calls in a day is $N(t)$, a poisson process with parameter $\lambda(x) = 1000x^4(1-x)$.

```{r 1a}
lambda1 <- function(x) {
    for (i in 1:length(x)){
        return (1000*x^4*(1-x))
    }
}
x <- seq(0,1,.01)
ggplot() + geom_line(mapping=aes(x=x, y=lambda1(x)))
```

(b) Note that the function $$\lambda(x) = 1000B(5,2)X$$ with $X \approx Beta(5,2)$.

Recall $\beta(5,2) = \frac{\Gamma(5)\Gamma(2)}{\Gamma(5+2)}$

```{r 1b}
set.seed(42)
MC <- 10000
# To find the value of m(tl)
m <- mean(lambda1(runif(MC, 0, 1)))
# For only one day
path1 <- function(m){
    k <- rpois(1, m)
    v <- 1000*beta(5,2)*rbeta(k, 5, 2)/m
    v <- sort(v)
    return(v)
}
t <- path1(m)
min <- function(t){
    m <- t[2]-t[1]
    for (i in 2:length(t)-1){
        if (t[i+1]-t[i] < m) m <- t[i+1]-t[i]
    }
    return(m)
}
min_t <- min(t)

process <- function(m){
    times <- path1(m)
    return(min(t))
}
min_times <- replicate(MC, process(m))

```

# Exercise 1
Computing the inverse we get $$F^{-1}(y) = \frac{2}{\sqrt[3]{1-y}},\,\,\! y>0$$
```{r 1}
set.seed(42)
fun <- function(t){
    for (i in 1:length(t)){
        if (t[i] > 2){
            f <- 1-(2/t)^3
        } else {
            f <- 0
        }
        return(f)
    }
}
inv <- function(){
    y <- runif(1, 0, 1)
    if (y > 0){
        return(2/(1-y)^(1/3))
    } else {
        return(0)
    }
}
MC <- 1000;
sims <- replicate(MC, inv())
ks.test(sims, fun)
ks.test(sims, 'ppareto', shape=3, location=2)
```
## Exercise 2
(a) Through mathematical computations, we have seen that for $M = \frac{4}{3}$, $f_S(t) \le M f_T(t)$.
(b) Therefore, $\frac{f_S(t)}{M f_T(t)} = \frac{2}{t}$. Following the algorithm in class.
```{r 2 b}
set.seed(42)
f.s.t <- function(t) 2/t
acceptance.rejection <- function(){
    f.t <- rpareto(1, location=2, shape=3)
    while (runif(1) > f.s.t(f.t)) f.t <- rpareto(1, location=2, shape=3)
    return(f.t)
}
```
(c)
```{r 2 c}
MC <- 10000
x <- replicate(MC, acceptance.rejection())
t.test(x, conf.level=.99)
```